---
title: "Mini Literate Analysis"
author: "Kathryn Tomasi"
format: html
---

## Introduction

### About this data

This data set contains North Pole Environmental Observatory Bottle Chemistry data. This data set includes diverse measurements of seawater chemistry along with supporting conductivity, temperature, and depth (CTD) data from the Arctic Ocean near the North Pole. It was retrieved on October 2, 2023 from [here.](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)


### Setup

```{r}
library(readr)
library(here)
```


### Read data

```{r}
#| warning: false
bg_chem <- read_csv("../data/BGchem2008data.csv")

bg_chem <- read_csv(here("data/BGchem2008data.csv"))
```

Exploring our data

```{r}
#| eval: false
colnames(bg_chem)
```
```{r}
str(bg_chem)
```
```{r}
summary(bg_chem)
```
```{r}
unique(bg_chem$Date)
```


## Analysis
### Calculate summary statistics

```{r}
nitrate <- mean(bg_chem$NO3)

nitrite <- mean(bg_chem$NO2)

amm <- mean(bg_chem$NH4)

phos <- mean(bg_chem$P)
```


### Calculate mean Redfield ratio

```{r}
ratio <- (nitrate + nitrite + amm / phos)
```

### Plot Ratio

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```


## Conclusion

The Redfield ratio of this data is `r round(ratio)`
